Unknown

### 2025-06-25 ms/ms.qmd
working on ms results

#### MARKER
<!-- Paragraph 6 (summarize main results) -->
Consistent with biophysical theory of CO$_2$ diffusion within leaves, AA > 0 for all accessions (Figure X). AA varied substantially between measurement light intensities, growth light intensities, and among accessions (point to model comparison evidence). Measured under high light intensity, AA was consistently greater for sun plants. The average AA among accessions in the shade treatment was `r filter(aa_mean_range, light_treatment == "low", light_intensity == "2000")$x`; however, the same accessions grown at high light intensity showed a mean AA of `r filter(aa_mean_range, light_treatment == "high", light_intensity == "2000")$x`. Contrary to the predictions of the assimilatory hypothesis, AA was greater in all accessions under low measurement light intensity for both sun and shade grown plants. The effect of low light on AA was more pronounced in the sun-grown plants, where AA was significantly greater under low measurement light intensity in `r aa_summary3["high"]` of `r n_acc` accessions compared to to `r aa_summary3["low"]` accessions for shade-grown plants. The overall average AA of shade and sun grown plants measured under low light intensity was `r filter(aa_mean_range, light_treatment == "low", light_intensity == "150")$x` and `r filter(aa_mean_range, light_treatment == "high", light_intensity == "150")$x`, respectively. There was a slight tendency for accessions from more closed habitats (greater PAI), but AA varied widely among accessions from open habitats (low PAI) regardless of growth and measurement light intensities (Figure X). The pattern of AA across wild tomatoes strongly supports the plasticity hypothesis, argues against the acclimatory hypothesis, and provides only weak support for the constitutive hypothesis.

@@

<!-- Paragraph 7 (summarize secondary results) -->
These results suggest that developmental plasticity, as opposed to acclimation or adaptation to open habitats, may explain the explain the long-standing observation taht amphistomatous leaves are more common in sunny habitats. Our results add to existing explanations by showing high light intensity *per se* does not increase the benefit of amphistomy, but rather than anatomical and biochemical change associated with higher light intensity modulate AA. We cannot yet attribute specific changes wrought by plasticity, but correlational evidence suggests that changes in leaf anatomy may be important. Species responded consistently by increasing stomatal density, especially on the adaxial surface, increasing stomatal density ratio. Sun leaves also had greater gs, Amax. However, these traits were not associated with change in AA (table/figures).

>>>>

<!-- Paragraph 6 (summarize main results) -->
Consistent with biophysical theory of CO$_2$ diffusion within leaves, AA > 0 for all accessions (Figure X). AA varied substantially between measurement light intensities, growth light intensities, and among accessions (point to model comparison evidence). Measured under high light intensity, AA was consistently greater for sun plants. The average AA among accessions in the shade treatment was `r filter(aa_mean_range, light_treatment == "low", light_intensity == "2000")$x`; however, the same accessions grown at high light intensity showed a mean AA of `r filter(aa_mean_range, light_treatment == "high", light_intensity == "2000")$x`. Contrary to the predictions of the assimilatory hypothesis, AA was greater in all accessions under low measurement light intensity for both sun and shade grown plants. The effect of low light on AA was more pronounced in the sun-grown plants, where AA was significantly greater under low measurement light intensity in `r aa_summary3["high"]` of `r n_acc` accessions compared to to `r aa_summary3["low"]` accessions for shade-grown plants. The overall average AA of shade and sun grown plants measured under low light intensity was `r filter(aa_mean_range, light_treatment == "low", light_intensity == "150")$x` and `r filter(aa_mean_range, light_treatment == "high", light_intensity == "150")$x`, respectively. There was a slight tendency for accessions from more closed habitats (greater PAI), but AA varied widely among accessions from open habitats (low PAI) regardless of growth and measurement light intensities (Figure X).

<!-- Paragraph 7 (summarize secondary results) -->
The pattern of AA across wild tomatoes strongly supports the plasticity hypothesis, argues against the acclimatory hypothesis, and provides only weak support for the constitutive hypothesis. Plastic changes in leaf thickness and/or packing density, summarized by the bulk leaf mass per area (LMA), likely mediate the effect of growth light intensity on AA. When LMA is included in the model, the effect of growth light intensity on AA is no longer significant based on model comparison (Table ? - need to check out what the interaction with measurement light intensity means). In contrast, LMA did not mediate indirect effects of native PAI on AA (Table ?). Many anatomical traits underlie LMA [@john_anatomical_2017] and future research will be needed identify which particular traits, such as leaf thickness or mesophyll porosity, are responsible for mediating AA.

#### MARKER
These results suggest that developmental plasticity, as opposed to acclimation or adaptation to open habitats, may explain the explain the long-standing observation taht amphistomatous leaves are more common in sunny habitats. Our results add to existing explanations by showing high light intensity *per se* does not increase the benefit of amphistomy, but rather than anatomical and biochemical change associated with higher light intensity modulate AA. We cannot yet attribute specific changes wrought by plasticity, but correlational evidence suggests that changes in leaf anatomy may be important. Species responded consistently by increasing stomatal density, especially on the adaxial surface, increasing stomatal density ratio. Sun leaves also had greater gs, Amax. However, these traits were not associated with change in AA (table/figures).

@@
